### 2016-02-22 ch02-qlogms/excitation-design.tex
Some comments from Tom + merge CH02

The information matrix tells us how good the individual parameter estimates are. However, if we are interested in estimating a transfer function $G(s,\theta)$ the quality of this estimate is of more importance. It is possible to find an expression for the variance of $\hat{G}(j\omega,\theta)$ using the Gauss approximation formula \citep{Ljung1999, Pintelon2012}. This gives
\begin{align}
\var{\hat{G}(j\omega)} &= \frac{\partial G^H(j\omega)}{\partial \theta}\IF^{-1}\frac{\partial G(j\omega)}{\partial \theta}.
\label{eq:excitation:varG}
\end{align}

Typically, the input design formulation depends on the true system parameters. This means that the optimal input depends on the \emph{unknown} system that one wants to identify. Furthermore, classical input designs, such as the D-optimal, typically give signals with sparse spectra. This means that only a few frequencies are excited. A sparse spectrum is bad from a robustness point of view; there is very limited information about what happens at the unexcited frequencies. Therefore, a signal design to be optimal for one system may result in very bad estimates if applied to another system. 
To increase the robustness of the design, one could take a small fraction $\delta$, from the optimal power and redistribute this power over the whole frequency band of interest.
From \eqref{eq:excitation:IF} and \eqref{eq:excitation:varG} we see that scaling the input power by $1-\delta$, gives the transfer function variance
\begin{align}
\var{\hat{G}(j\omega)}\frac{1}{1-\delta}.
\end{align}
This is a small loss while the increase in robustness can be significant if the added signal is robust. Such a robust signal is studied in the following parts of the chapter.

@@

%\subsection{Using prior system information}
%Consider a system $G(s,\theta)$ where some initial guess of the parameter vector,
% say $\hat{\theta}_0 \in \mathbb{R}^n$ exists. A D-optimal design can then be
% found based $\hat{\theta}_0$. The resulting determinant is
%\begin{align}
%  \left[\det\IF^\star\right]^{-1}
%    &= \min_{\phi_u(\omega)}
%             \left[\det\IF\right]^{-1}
%\end{align}
%
%The optimal input can be realized as sum of $n$ sinusoids.
%Such an input is typically not robust to the initial guess since \TODO{\ldots}.
%As a solution to this, the optimal input is modified so that a fraction of the
%power in the optimal design is taken away and redistributed over all frequencies.
%This ensures that some excitation power is available also at other frequencies.
%
%If $\lambda_i$ is the fraction of the input power at frequency $\wn[i]$,
%then the determinant of $\IF^\star$ can be expressed as \Cite{Goodwin}
%\begin{align}
%  \det\IF^\star &= \prod_{i = 1}^{n}
%                      \lambda_i^2
%                      \alpha(\wn[i]).
%\end{align}
%The optimal input spectrum is
%\begin{align}
%  \phi_u^\star(\omega) &= \sum_{i=1}^{n}
%                            \lambda_i^2
%                            \delta(\omega,\wn[i])
%\end{align}
%If a fraction, say $\eta$, of the power is taken from each frequency and used
%at other frequencies, the input spectrum can be written as
%\begin{align}
%  \phi_u(\omega) &= (1-\eta)
%                      \sum_{i=1}^{n}
%                        \lambda_i^2 
%                        \delta(\omega,\wn[i])
%                    + \phi_u^{\robust}(\omega),\\
%                 &= (1-\eta) \phi_u^\star    (\omega)
%                    +        \phi_u^{\robust}(\omega).
%\end{align}
%The term $ \phi_u^{\robust}$ is the part that introduces robustness and
%must be such that
%\begin{align}
%  \int_0^{\pi} \phi_u^{\robust}(\omega) \dd{\omega} &= n \eta.
%\end{align}
%The spectrum $\phi_u^{\robust}(\omega)$ can for example be chosen as
%a discrete spectrum, e.g. from a multi-sine signal, or as
%a continuous spectrum. The resulting information matrix for the robust input is hence
%\begin{align}
%  \IF &= (1-\eta)\IF^\star + \IF^\robust
%\end{align}
%The determinant is therefore
%\begin{multline}
%\det\left[(1-\eta)\IF^\star + \IF^\robust\right]
%  \geq \det\left[(1-\eta)\IF^\star\right] + \det\left[\IF^\robust\right] \\
%  =  (1-\eta)^n\det\left[\IF^\star\right] + \det\left[\IF^\robust\right]
%  \text{.}
%\end{multline}
%The inequality holds since the information matrix is a positive, semi-definite
%matrix. Hence the largest possible loss in the determinant due to the robustness
% is given by the factor $(1-\eta)^n$.
%\TODO{This bound is often very low, maybe better one should be found!}

@@

Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2012}] given by
\begin{align}
  \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\delta(\omega,\wn[K])},
\end{align}
where $\omega_{k}/\omega_{k+1} \approx \alpha$.
Hence, the number of spectral lines inside \wdB is 
\begin{align}
  \floor{
    \log_{\alpha} \frac{1+\damping}{1-\damping} 
  }
  \text{,}
\end{align}
where $\floor{\placeholder}$ is the nearest integer rounded towards $0$. The number of spectral lines is independent of $\wn$.

@@

@@

Direct calculations for $\damping < 1$ give
\begin{align}
\label{eq:excitation:auto}
\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega = \frac{\pi}{16\wn[1]^2}\damping^{-3} + \mathcal{O}(\damping^{-1}).
\end{align}
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
% \begin{align}
% &\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right| \leq\label{eq:excitation:cross1} \\
% &\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega \leq \\
% &\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega.\label{eq:excitation:cross}
% \end{align}
\begin{gather}
\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right|
\qquad \qquad \qquad \qquad \qquad
\label{eq:excitation:cross1} \\
\leq
\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega \text{.}\label{eq:excitation:cross}
\end{gather}
To proceed, consider
\begin{align}
\frac{\partial G_1}{\partial \wn[1]} 
  &= \frac{2\damping j\omega\left[1+\frac{j\omega}{\wn[1]\damping}\right]}
  {\wn[1]^2\left[1+2\frac{\damping j\omega}{\wn[1]} + \left(\frac{j\omega}{\wn[1]}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\left\{0, \damping\wn[1]\right\}$. 
Hence $\left|\frac{\partial G_1}{\partial \wn[1]} \right|$ has asymptotic slope 20\unit{dB/\decade} for low frequencies, at $j\omega=\damping\wn[1]$ the slope increases to 40\unit{dB/\decade}. 
There are four poles which give a $-40$\unit{dB/\decade} roll-off after the peak. 
The shape of $\frac{\partial G_1}{\partial \wn[1]} $ is illustrated in \figref{fig:excitation:dG}.

@@

\begin{figure}
\centering
\setlength\figurewidth{0.68\columnwidth}
\setlength\figureheight{0.68\figurewidth}
\input{\thisDir/figs/rolloff.tikz}
\caption{The shape of $\frac{\partial G_1}{\partial \wn[1]}$. The roll-off around the peak is significant and depends on the damping of the system.}
\label{fig:excitation:dG}
\end{figure}

@@

In \figref{fig:excitation:coupling} the largest element in $\partial G_{1,2}$ is plotted for different dampings and separations. It is clear that highly resonant systems have a large degree of decoupling and that resonances that are well separated increases the decoupling. The diagonal blocks of the normalized $\IF$ in all cases have some elements close to 1. In this example, the $\damping$ is set to be the same in both sub-systems. Using different dampings in the sub-systems would, however, not change the argument, the systems with the lowest damping will limit the separation.

@@

  The model is parametrized in the vector $\theta = \left[b_0 \,\cdots\, b_{n_b-1} \; a_0 \,\cdots\, a_{n_a-1} \right]^{\T}$ and  is evaluated at different frequencies $\Omega$ in the complex plane.
  For a continuous time model $\Omega = s$, while for the discrete time case $\Omega = z$.

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion.
  \TODO{This is actually wrong, rather, it should be $\frac{as-b}{cs^2 + ds + 1}$. However, the difference in information is rather small around the actual resonance. See script in NnF folder.}
\TODO{write out in partial fraction expansion, use script to show that in practice, only a few parameters matter. This will be well-approximated by considering only the second order system of the paper.}

  Within these sub-systems, we can distinguish first-order subsystems and second-order subsystems, the latter of which describe resonant phenomena.
  Since such resonances are often a limiting factor, we first focus on these resonant sub-systems.
  Afterwards, it will become obvious that

@@

  This last example illustrates that a least common multiple of real numbers does not always exist.
  Consequently, the generalized multisine of \defref{def:generalized-MS} is not guaranteed to be periodic.
  In actual implementations with limited numerical precision, this gives rise to signals that are periodic but whose period can be too long to be practically viable for periodic measurements.
  Consequently, it is undesirable to use generalized multisines that have an overly long period.

@@

  When designing an excitation signal, it is common only to specify a (large) frequency band
  by its boundary frequencies $f_{\min}$ and $f_{\max}$.
  For a given $\alpha$ it is easy to determine the number of excited lines
  within the frequency range of interest:
    \begin{equation}
      F = \floor{
                  \frac{\log{f_{\max}} - \log{f_{\min}}}
                       {\log{\alpha}}
                }
    \label{eq:logms:F}
    \end{equation}
  where $\floor{\bullet}$ denotes rounding towards zero.
  In the Section \ref{sec:optimAlpha}, the choice of a suitable $\alpha$ is elaborated.

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} or the \code{logspace} function in \MATLAB it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$).
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

   Obviously, the same argument can be held for any set of frequency bands with constant relative bandwidth (e.g. octave $\kappa = 2$).
   I.e. the same number of frequency components is present per decade, such that the power in each decade is identical when a constant amplitude spectrum ($A_k = A_{\mathrm{in}}$) is used.

  To return to the FRF, in Section~\ref{sec:optimAlpha} we will show that for a dense logarithmic frequency grid, any second order system with damping $\damping$ and a resonance within the bulk of the frequency grid, will receive the same number of frequency lines in its $3 \unit{dB}$ bandwidth regardless of the actual value of the resonance frequency.
  As most of the information of such systems is obtained from the measurements at frequencies inside the $3 \unit{dB}$
  bandwidth, one expects an equal variance for each system, as will be shown in Section~\ref{sec:bestFrequencyResolution}.

  For that reason, a log multisine is a good candidate for our purposes.
  However, it is more involved both to generate and to analyze such a signal, as the excited frequencies don't lie on a commensurate frequency grid.
  Especially for periodic measurements, this is a major downside since the period of such signal does not always exist (i.e. this leads to infinite measurement times).
  It is therefore not practical to use a perfectly logarithmically spaced frequency grid as the \gls{DFT} or \gls{CZT}~\citep{Rabiner2004} explicitly rely on a commensurate frequency grid.

@@

  \begin{definition}
  For the implementation of a quasi-logarithmic frequency grid,  the approximate equality \eqref{eq:qlogGrid} does not define the grid unambiguously.
  The quasi-logarithmic grids depicted in this chapter are constructed using the relationship
  \begin{align}
       \label{eq:excitation:qlog:gridDef}
    f_k & = \alpha_k(N_k) f_{k-1}  
        \quad
            \forall k \in \set{1,\ldots,F}
\end{align}
with
\begin{align}
     \label{eq:excitation:qlog:roundAlpha}
      \alpha_k(N_k) & \isdef 
          \round{   \frac{\alpha^{N_k} f_{k-1}}{\Delta f}}    
           \frac{\Delta f}{f_{k-1}}
      \approx \alpha \text{, and}\\
         \label{eq:excitation:qlog:alphaPowerDef}
          N_k & \isdef 
            \min 
            \set{n | n \in \NaturalNumbers, \alpha_k(n) > 1}
     \text{.}
  \end{align}
  \end{definition}
  In the equation above, $\round{\placeholder}$ denotes rounding towards the nearest integer according to the \IEEEfloat standards and as implemented in the \MATLAB function \code{round}.
  This satisfies the approximate definition.

  \begin{remark}
  The use of different rounding strategies --- e.g. the \code{ceil} function $\ceil{}$  and \code{floor} function $\floor{}$ --- in \eqref{eq:excitation:qlog:roundAlpha}, allows for slightly different guarantees on the effective $\alpha_k$ attained by the grid.
  By denoting $\alpha_k^{\placeholder}$ as the effective frequency ratio for a rounding strategy $\placeholder$  in that equation, one can trivially see that the following relationships hold:
  \begin{equation}
  \alpha_k^{\floor{}} \leq \alpha \leq  \alpha_k^{\ceil{}}
  \qquad \text{and} \qquad
  \alpha_k^{\floor{}} \leq \alpha_k^{\round{}} \leq  \alpha_k^{\ceil{}}
  \qquad
  \forall k \in \NaturalNumbersWithoutZero
  \text{.}
  \end{equation}
  This means that rounding using the \code{floor} operator favors to strictly attain the specified frequency bin density at the cost of a few more excited lines (and hence less allottable power per line).
  The \code{ceil} function produces a grid that has more power per line at the cost of a reduced bin density.
  On the other hand, using \code{round} as in \eqref{eq:excitation:qlog:roundAlpha}, provides a grid that is somewhere in-between both extremes.
  \end{remark}

>>>>

The information matrix tells us how good the individual parameter estimates are. However, if we are interested in estimating a transfer function $G(s,\theta)$ the quality of this estimate is of more importance. It is possible to find an expression for the variance of $\hat{G}(j\omega,\theta)$ using the Gauss approximation formula \citep{Ljung1999, Pintelon2012}. This gives
\begin{align}
\var{\hat{G}(j\omega)} &\approx \frac{\partial G^H(j\omega)}{\partial \theta}\IF^{-1}\frac{\partial G(j\omega)}{\partial \theta}.
\label{eq:excitation:varG}
\end{align}

Typically, the input design formulation depends on the true system parameters. This means that the optimal input depends on the \emph{unknown} system that one wants to identify. 
Furthermore, classical input designs, such as the D-optimal design, typically give signals with sparse spectra. 
This means that only a few frequencies are excited. 
A sparse spectrum is bad from a robustness point of view; there is very limited information about what happens at the unexcited frequencies. 
Therefore, a signal design to be optimal for one system may result in very bad estimates if applied to another system. 
To increase the robustness of the design, one could take a small fraction $\delta$, from the optimal power and redistribute this power over the whole frequency band of interest.
From \eqref{eq:excitation:IF} and \eqref{eq:excitation:varG} we see that scaling the input power by $1-\delta$, gives the transfer function variance
\begin{align}
\var{\hat{G}(j\omega)}\frac{1}{1-\delta}.
\end{align}
This is a small loss while the increase in robustness can be significant if the added signal is robust. Such a robust signal is studied in the following parts of the chapter.

@@

% \TODO{begin}
% \newcommand{\robust}{\mathrm{robust}\}
% \subsection{Using prior system information}
% Consider a system $G(s,\theta)$ where some initial guess of the parameter vector,
% say $\hat{\theta}_0 \in \mathbb{R}^n$ exists. A D-optimal design can then be
% found based $\hat{\theta}_0$. The resulting determinant is
% \begin{align}
%  \left[\det\IF^\star\right]^{-1}
%    &= \min_{\phi_u(\omega)}
%             \left[\det\IF\right]^{-1}
% \end{align}

% The optimal input can be realized as sum of $n$ sinusoids.
% Such an input is typically not robust to the initial guess since \TODO{\ldots}.
% As a solution to this, the optimal input is modified so that a fraction of the
% power in the optimal design is taken away and redistributed over all frequencies.
% This ensures that some excitation power is available also at other frequencies.

% If $\lambda_i$ is the fraction of the input power at frequency $\wn[i]$,
% then the determinant of $\IF^\star$ can be expressed as \Cite{Goodwin}
% \begin{align}
%  \det\IF^\star &= \prod_{i = 1}^{n}
%                      \lambda_i^2
%                      \alpha(\wn[i]).
% \end{align}
% The optimal input spectrum is
% \begin{align}
%  \phi_u^\star(\omega) &= \sum_{i=1}^{n}
%                            \lambda_i^2
%                            \delta(\omega,\wn[i])
% \end{align}
% If a fraction, say $\eta$, of the power is taken from each frequency and used
% at other frequencies, the input spectrum can be written as
% \begin{align}
%  \phi_u(\omega) &= (1-\eta)
%                      \sum_{i=1}^{n}
%                        \lambda_i^2 
%                        \delta(\omega,\wn[i])
%                    + \phi_u^{\robust}(\omega),\\
%                 &= (1-\eta) \phi_u^\star    (\omega)
%                    +        \phi_u^{\robust}(\omega).
% \end{align}
% The term $ \phi_u^{\robust}$ is the part that introduces robustness and
% must be such that
% \begin{align}
%  \int_0^{\pi} \phi_u^{\robust}(\omega) \dd{\omega} &= n \eta.
% \end{align}
% The spectrum $\phi_u^{\robust}(\omega)$ can for example be chosen as
% a discrete spectrum, e.g. from a multi-sine signal, or as
% a continuous spectrum. The resulting information matrix for the robust input is hence
% \begin{align}
%  \IF &= (1-\eta)\IF^\star + \IF^\robust
% \end{align}
% The determinant is therefore
% \begin{multline}
% \det\left[(1-\eta)\IF^\star + \IF^\robust\right]
%  \geq \det\left[(1-\eta)\IF^\star\right] + \det\left[\IF^\robust\right] \\
%  =  (1-\eta)^n\det\left[\IF^\star\right] + \det\left[\IF^\robust\right]
%  \text{.}
% \end{multline}
% The inequality holds since the information matrix is a positive, semi-definite
% matrix. Hence the largest possible loss in the determinant due to the robustness
% is given by the factor $(1-\eta)^n$.
% \TODO{This bound is often very low, maybe better one should be found!}
% \TODO{end}

@@

Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2012}] given by
\begin{align}
  \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\dirac{\omega - \omega_k}},
\end{align}
where $\omega_{k}/\omega_{k+1} \approx \alpha$ and $\dirac{\omega}$ denotes the Dirac distribution.
Hence, the number of spectral lines inside \wdB is 
\begin{align}
  \floor{
    \log_{\alpha} \frac{1+\damping}{1-\damping} 
  }
  \text{,}
\end{align}
where $\floor{\placeholder}$ is the nearest integer rounded towards $0$. The number of spectral lines is independent of $\wn$.

@@

\subsection{Generic system}
%TODO: explain generic LTI system
Consider a second-order subsystem
\begin{equation}
  G_2(s, \theta) = \frac{K \wn^{2 \multiplicity} \left( s + \wz \right)}{\wz \left(s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity}}
\end{equation}
parametrized in $\theta =  \left[K, \wn, \wz, \damping \right]^{\T}$.
To determine the Fisher Information matrix, the derivative $\PartialDerivative{G(s,\theta)}{\theta}$ is required; its components are given by:
\begin{align}
  \PartialDerivative{G_2}{K} 
    &= 
  \frac{\wn^{2\multiplicity}}{\wz}
  \frac{s + \wz}{\left(s^2 + 2 \damping \wn s + \wn^2\right)^{\multiplicity}}
  \\
  \PartialDerivative{G_2}{\wz} 
    &= 
    \frac{-K \wn^{2\multiplicity}}
              {\wz} 
    \frac{s}
              {\left(s^2 + 2 \damping \wn s + \wn^2\right)^{\multiplicity}} 
  \\
  \PartialDerivative{G_2}{\wn} 
  &= 
  \frac{2 K \wn^{2\multiplicity-1} \multiplicity}
            {\wz} 
  \frac{\left( s + \wz \right) \left( s+ \damping \wn \right)}
            {\left( s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity + 1}}
  \\
  \PartialDerivative{G_2}{\damping}
  &= 
  \frac{-2 K \wn^{2\multiplicity + 1} \multiplicity}
            {\wz}
  \frac{s \left( s + \wz \right)}
             {\left( s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity + 1}}
  \text{.}
\end{align}

@@

Direct calculations for $\damping < 1$ give
\begin{align}
\label{eq:excitation:auto}
\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega = \frac{\pi}{16\wn[1]^2}\damping^{-3} + \mathcal{O}(\damping^{-1}).
\end{align}
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
% \begin{align}
% &\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right| \leq\label{eq:excitation:cross1} \\
% &\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega \leq \\
% &\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega.\label{eq:excitation:cross}
% \end{align}
\begin{gather}
\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right|
\qquad \qquad \qquad \qquad \qquad
\label{eq:excitation:cross1} \\
\leq
\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega \text{.}\label{eq:excitation:cross}
\end{gather}
To proceed, consider
\begin{align}
\frac{\partial G_1}{\partial \wn[1]} 
  &= \frac{2\damping j\omega\left[1+\frac{j\omega}{\wn[1]\damping}\right]}
  {\wn[1]^2\left[1+2\frac{\damping j\omega}{\wn[1]} + \left(\frac{j\omega}{\wn[1]}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\left\{0, \damping\wn[1]\right\}$. 
Hence $\left|\frac{\partial G_1}{\partial \wn[1]} \right|$ has asymptotic slope 20\unit{dB/\decade} for low frequencies, at $j\omega=\damping\wn[1]$ the slope increases to 40\unit{dB/\decade}. 
There are four poles which give a $-40$\unit{dB/\decade} roll-off after the peak. 
The shape of $\frac{\partial G_1}{\partial \wn[1]} $ is illustrated in \figref{fig:excitation:derivatives}.

@@

@@

In \figref{fig:excitation:coupling} the largest element in $\partial G_{1,2}$ is plotted for different dampings and separations. It is clear that highly resonant systems have a large degree of decoupling and that resonances that are well-separated increases the decoupling. 
The diagonal blocks of the normalized $\IF$ in all cases have some elements close to $1$. 
In this example, the $\damping$ is set to be the same in both sub-systems. 
Using different dampings in the sub-systems would, however, not change the argument, the systems with the lowest damping will limit the separation.

@@

  The model is parametrized by the vector $\theta = \left[b_0 \,\cdots\, b_{n_b-1} \; a_0 \,\cdots\, a_{n_a-1} \right]^{\T}$ and  is evaluated at different frequencies $\Omega$ in the complex plane.
  For a continuous time model $\Omega = s$, while for the discrete time case $\Omega = z$.

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion:
  \TODO{move into main text}
Remember that the rational transfer function of any strictly proper \gls{LTI} system $G$ can be decomposed in its partial fraction form:
\begin{align}
  G(s) &= 
    \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\sum\limits_{i=0}^{\numel{A}} a_i s^i} 
        =    
   \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\prod\limits_{i=1}^{I_1} \left( s - \pole[i] \right)^{M_{1,i}} 
               \prod\limits_{i=1}^{I_2} \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{M_{2,i}}}
        \\
        &  = 
        \sum_{i=1}^{I_1} 
          \sum_{\multiplicity=1}^{M_{1,i}} 
            \frac{\dc[i]}
                      { \left( s - \pole[i] \right)^{\multiplicity}}
            + 
        \sum_{i=1}^{I_2} 
          \sum_{\multiplicity=1}^{M_{2,i}}
           \frac{\dc[i,\multiplicity] \wn[i]^{2\multiplicity} \left( s + \wz[i] \right)}
                     {\wz[i] \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{\multiplicity}}
        \label{eq:excitation:PFE}
\end{align}
where $I_1$ denotes the number of real-valued poles, $I_2$ denotes the number of complex conjugated pole pairs.
$M_{1,i}$ and $M_{2,i}$ denotes the multiplicity of the corresponding poles, such that $\numel{A} = \sum_i^{I_1} M_{1,i} + 2 \sum_{i}^{I_2} M_{2,i}$.
Each term in \eqref{eq:excitation:PFE} represents a \term{sub-system} of the system $G(s)$.  
  Within these sub-systems, we can distinguish first-order subsystems and second-order subsystems, the latter of which describe resonant phenomena.
  Since such resonances are often a limiting factor, we first focus on these resonant sub-systems.

@@

  This last example illustrates that a least common multiple of real numbers is not guaranteed to exist.
  Consequently, the generalized multisine of \defref{def:generalized-MS} is not guaranteed to be periodic.
  In actual implementations with limited numerical precision, this gives rise to signals that are periodic but whose period can be too long to be practically viable for periodic measurements.
  Consequently, it is undesirable to use generalized multisines that have an overly long period.

@@

  When designing an excitation signal, it is common only to specify a (large) frequency band
  by its boundary frequencies $f_{\min}$ and $f_{\max}$.
  For a given $\alpha$ it is easy to determine the number of excited lines
  within the frequency range of interest:
    \begin{equation}
      F = \floor{
                  \frac{\log{f_{\max}} - \log{f_{\min}}}
                       {\log{\alpha}}
                }
    \label{eq:logms:F}
    \end{equation}
  where $\floor{\bullet}$ denotes rounding towards the nearest smaller integer value.
  In the Section \ref{sec:optimAlpha}, the choice of a suitable $\alpha$ is elaborated.

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}
  \begin{example}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}

   From \propref{prop:excitation:log:relBW}, it can be seen that for any set of frequency bands with a constant relative bandwidth (e.g. decades~($\kappa=10$), octaves~($\kappa=2$),\ldots ), a logarithmic  frequency grid contains a constant number of lines per frequency band.
   Consequently, the power in each relative bandwidth is identical when a constant amplitude spectrum ($A_k = A_{\mathrm{in}}$) is used.

  To return to the \gls{FRF}, in \secref{sec:optimAlpha} we will show that for a dense logarithmic frequency grid, any second order system with damping $\damping$ and a resonance within the bulk of the frequency grid, will receive the same number of frequency lines in its $3 \unit{dB}$ bandwidth regardless of the actual value of the resonance frequency.
  As most of the information of such systems is obtained from the measurements at frequencies inside the $3 \unit{dB}$
  bandwidth, one expects an equal variance for each system, as will be shown in \secref{sec:bestFrequencyResolution}.

  For that reason, a logarithmic multisine is a good candidate for our purposes.
  However, it is more involved both to generate and to analyze such a signal, as the excited frequencies don't lie on a commensurate frequency grid.
  Especially for periodic measurements, this is a major downside since the period of such signal does not always exist (i.e. this leads to infinite measurement times).
  It is therefore not practical to use a perfectly logarithmically spaced frequency grid as the \gls{DFT} or \gls{CZT}~\citep{Rabiner2004} explicitly rely on a commensurate frequency grid.

@@

  \begin{definition}
  For the implementation of a quasi-logarithmic frequency grid,  the approximate equality \eqref{eq:qlogGrid} does not define the grid unambiguously.
  The quasi-logarithmic grids depicted in this chapter are constructed using the relationship
  \begin{align}
       \label{eq:excitation:qlog:gridDef}
    f_k & = \alpha_k(N_k) f_{k-1}  
        \quad
            \forall k \in \set{1,\ldots,F}
\end{align}
with
\begin{align}
     \label{eq:excitation:qlog:roundAlpha}
      \alpha_k(N_k) & \isdef 
          \round{   \frac{\alpha^{N_k} f_{k-1}}{\Delta f}}    
           \frac{\Delta f}{f_{k-1}}
      \approx \alpha \text{, and}\\
         \label{eq:excitation:qlog:alphaPowerDef}
          N_k & \isdef 
            \min 
            \set{n | n \in \NaturalNumbers, \alpha_k(n) > 1}
     \text{.}
  \end{align}
  \end{definition}
  In the equation above, $\round{\placeholder}$ denotes rounding towards the nearest integer according to the \IEEEfloat standards and as implemented in the \MATLAB function \mcode{round}.
  This satisfies the approximate definition.

  \begin{remark}
  The use of different rounding strategies --- e.g. the \mcode{ceil} function $\ceil{}$  and \mcode{floor} function $\floor{}$ --- in \eqref{eq:excitation:qlog:roundAlpha}, allows for slightly different guarantees on the effective $\alpha_k$ attained by the grid.
  By denoting $\alpha_k^{\placeholder}$ as the effective frequency ratio for a rounding strategy $\placeholder$  in that equation, one can trivially see that the following relationships hold:
  \begin{equation}
  \alpha_k^{\floor{}} \leq \alpha \leq  \alpha_k^{\ceil{}}
  \qquad \text{and} \qquad
  \alpha_k^{\floor{}} \leq \alpha_k^{\round{}} \leq  \alpha_k^{\ceil{}}
  \qquad
  \forall k \in \NaturalNumbersWithoutZero
  \text{.}
  \end{equation}
  This means that rounding using the \mcode{floor} operator favors to strictly attain the specified frequency bin density at the cost of a few more excited lines (and hence less allottable power per line).
  The \mcode{ceil} function produces a grid that has more power per line at the cost of a reduced bin density.
  On the other hand, using \mcode{round} as in \eqref{eq:excitation:qlog:roundAlpha}, provides a grid that is somewhere in-between both extremes.
  \end{remark}
